Closing the gap between statistical and scientific workflows for improved forecasts in ecology

### 2025-03-24 forecastflows.tex
Just some random notes from thinking this weekend.

\title{Closing the gap between statistical and scientific workflows for improved forecasts in ecology } 
\date{\today}
\author{Victor Van der Meersch, J. Regetz, T. J. Davies \& EM Wolkovich}\\
 \maketitle

@@

Miscellaneous notes/points without a home
\begin{enumerate}
\item Ecologists need to race the same data to make progress for trends and for forecasting (point to make at end of paper maybe? And what is the workflow for this?) ... though LPI is used a lot, perhaps it is  sign that ecology is ready to start racing the same data, but then we need `analysis-ready' data so we're not all slightly differently cleaning the data etc..
\item Machine learning threatens utility of PBMs
\item We need more uncertainty propagation for trends and forecasting (uncertainty is esp. ignored in PBMs)
\item This workflow should lead to less model comparison (AIC, stepwise)
\item This workflow works for machine learning!
\item PBM: workflow should require estimating all parameters together; data simulation should reduce parameter number and highlight non-biol results
\item CUrrent trend workflow: SHould be research questi
\end{enumerate}

>>>>

\title{Closing the gap between statistical and scientific workflows for improved forecasts in ecology } 
\date{\today}
\author{Victor Van der Meersch, J. Regetz, T. J. Davies$^*$ \& EM Wolkovich}\\
 \maketitle
$^*$ Says he is happy to help and give friendly review, but not sure he will reach level of co-author.

@@

Miscellaneous notes/points without a home
\begin{enumerate}
\item Ecologists need to race the same data to make progress for trends and for forecasting (point to make at end of paper maybe? And what is the workflow for this?) ... though LPI is used a lot, perhaps it is  sign that ecology is ready to start racing the same data, but then we need `analysis-ready' data so we're not all slightly differently cleaning the data etc..
\item Machine learning threatens utility of PBMs
\item We need more uncertainty propagation for trends and forecasting (uncertainty is esp. ignored in PBMs)
\item This workflow should lead to less model comparison (AIC, stepwise)
\item This workflow works for machine learning!
\item PBM: workflow should require estimating all parameters together; data simulation should reduce parameter number and highlight non-biol results
\item Current trend workflow: Should be research question focused?
\end{enumerate}

Miscellaneous notes/points from thinking over 22-23 March weekend ...
\begin{enumerate}
\item Workflows emphasize there is no easy fix to better science or better stats
\item Maybe do a retrodictive check example with LPI? Including showing how even the Freckleton work could do more? 
\item What's the pathway from model comparison of many covariates to something else? Sometimes it seems like it's just prediction shoved into a mechanistic study, but if the goal is mechanism, there needs to be more work that either models these covariates together in a useful way (sort of approaching process-models!) or gets down to the fewer extremely relevant ones. 
\item When do we need to open up the black box? Something about we need better training for what science is and our goals; machine learning is not often helping with advancing \emph{science}
\item Scott Collins point that forecasting is not science (he claims a bunch of economists, weather folks etc. came to an ecological forecasting meeting and tried to explain that forecasting is an outcome of science, but it's not something you do science ON)
\end{enumerate}
